### 2016-07-22 correctness.tex
cheers

\section{Security Notions for Hash-Based Filters}

\heading{FP-security of hash-based filters.}
%\tsnote{NY give it in the private hash setting (sort of); but, intuitively, it should be hard to find false-positives even if you know that MD5 or SHA-x are the hash functions.  Not totally sure how to capture this intuition nicely.}
%
%\jnote{Not sure I agree. First of all, if $\epsilon$ is not negligible then a random element is a false positive with noticeable probability. Beyond that, I think that intuitively it makes sense that if you want false positives to be hard to find then you need to make the hash functions secret.}
%
%\tsnote{I don't think it makes sense to talk about asymptotics or negligibility, for exactly the reason you state. I think the key point here is: what is ``hard''? The designer decides what FP rate is acceptable for a given application, a given set of efficiency constraints, the side-effects of a FP, etc.  For a fixed FP target, she optimizes the filter size and/or the number of hash functions for her setting. Whether the target FP rate is 0.01 or $2^{-80}$, we should be able to say (in the ROM, at least) that it will take this-many hash computations, and this many on-line set-membership queries, to find the first FP.  In the standard model, with public hash functions, it is clearly more difficult to make formal claims.  Even so, in practice I expect it would be quite expensive to find a string~$x$ such that $(h_1(x),h_2(x),\ldots,h_k(x))$ falls in a given set of tuples when $h_i(x)=\mathrm{truncate}(\mathrm{SHA256}(\langle i \rangle \concat x), m)$, say. (Even more so if my setting admits relatively expensive hash computations, e.g. SCRYPT.) BTW, your suggestion to have a FP notion with secret~$\mathcal{S}$ is interesting here. }
%
The traditional notion of soundness for a Bloom filter captures the difficulty of finding false positives when the attacker is non-adaptive. Following NY, we consider soundness against adaptive adversaries. But we further generalize to allow for settings in which the attacker may have some a priori knowledge about the contents of~$\calS$. (The Niedermaier et al.~\cite{xxx} attacks, for example, exploit knowledge about the distribution of German last names.) In particular, our notions take as input an explicit distribution $\distr{U}{n}$ over~$[U]^n$.

>>>>

\section{Correctness Against Adaptive Adversaries}
As highlighted by Naor and Yogev, Equation~(\ref{eqn:correctness}) is a non-adaptive version of correctness;
we are interested in defining an adaptive (i.e., adversarial) notion
of correctness.   Figure~\ref{fig:correctness} describes a security
experiment that captures such a notion.

First, a set~$S$ is chosen according to some
distribution~$\distr{\calS}{}$ over the collection of multisets
$\mathcal{S}$ associated to~$\Pi=(\Rep,\Qry)$, and then
$\Rep(S)$ is run to generate $(M,\aux)$.
%(Unless stated otherwise, we assume security holds for all distributions and thus for all sets $S$.)
Adversary~$A$ is given~$\calS$ and $\aux$~as input; it is
provided an oracle~$\TestOracle$ that, on input a query~$q \in \calQ$,
computes (and returns) the result of $\Qry(M,\aux,q)$, and increments
a counter if this result is in error.  The adversary wins if it
manages to cause at least~$r>0$ such errors among its queries.

The notion is lifted to the random-oracle model by providing
$\Rep,\Qry$ and the adversary~$A$ with oracle access to the
random-oracle(s).

\begin{figure}[htp]
\centering
\fpage{.6}{
\hpagess{.6}{.35}
{
\experimentv{$\ExpCorrect{\setprim,r}{\distr{\calS}{},A}$}\\
$S \getsr \distr{\calS}{}$\\
$\mathrm{err}\gets 0$\\
$(M,\aux) \getsr \Rep(S)$\\
$z \getsr A^{\TestOracle}(S,\aux)$\\
if $\mathrm{err} < r$ then Return 0\\
Return 1
}
%
{
\oracle{$\TestOracle(q)$}\\
$a \gets \Qry(M,\aux,q)$\\
if $a \neq q(S)$ then \\
\nudge $\mathrm{err}\gets\mathrm{err}+1$\\
Return~$a$
}
}
\caption{Correctness of~$\Pi$ against (adaptive) adversary~$A$, when
  the target multiset~$S$ is sampled from~$\calS$ according to distribution~$\distr{\calS}{}$.} 
\label{fig:correctness}
\end{figure}

In this experiment we track the time-complexity (relative to some
implied model of computation) and query-complexity (i.e., number of
queries) of the adversary~$A$.  Roughly, we say that a
set-representation~$\Pi$ is $(t,q,r,\epsilon')$-correct if the
probability that any adversary with time-complexity~$t$ and
query-complexity~$q$ causes~$r$ errors with probability at
most~$\epsilon'$.  \tsnote{Maybe not... revisit.}

\def\bin{{\sf Bin}}
\heading{Defining adversarial advantage. } 
For any scheme~$\Pi$ there is a trivial,
\emph{non-adaptive} attack against this that wins a noticible fraction of the time.  In
particular, let $q_1,q_2,\ldots,q_T$ be uniform samples from~$\calQ$.
%The expected number of random queries needed to
%find a error is $1/\epsilon$ tries; thus, an attacker making $O(t/\epsilon)$ queries succeeds with constant probability (almost always).
%Actually, it seems worthwhile to formalize this. 
Let $X_i$ be a random variable indicating the event $Qry(M,\aux,q_i)\neq q_i(S)$.  
If~$\Pi$ is $\epsilon$-correct, then the probability that $X_i=1$ is at most~$\epsilon$.
And if the experiment requires the adversary to induce~$r$ errors, the attack wins iff
$N \stackrel{{\rm def}}{=} \sum_{i=1}^T X_i \geq r$.  Modelling the $X_i$ as independent Bernoulli trials with
probability~$\epsilon$ gives $N$ a binomial distribution.  \fixme{Something feels wrong with this...}  
Writing $\bin(T, \epsilon; r)$ for the probability that $N \geq r$,
we conclude that the \emph{best} correctness we can hope for against adversaries making~$T$ queries is  $\bin(T,\epsilon; t))$. %\jnote{Need to find different terms to use for a query $q \in \mathcal{Q}$
%and a query that the attacker makes to its oracle.}
\tsnote{Viewed another way, we could let~$T$ be a negative binomial RV, and ask how many queries are needed (w.h.p) for $N \geq t$.}

\jnote{One option is to define the adversary's \emph{advantage} as the difference between
its success probability and the success probability of the above,
trivial attack. Just a thought.}
\tsnote{Or, perhaps by recasting security
  as distinguishing between $\Qry(M,\aux,\cdot)$ and an oracle that
  returns 0,1 according to Bernoulli with paramter~$\epsilon$.}

\heading{Comparison to Naor-Yogev definition.}
The style of the Naor-Yogev (NY) definition is slightly different. \jnote{Even if we ignore
all the other issues, chief among them the inconsistent way it treats the hash functions.}
Lifting their definition to our setting, the experiment is similar but the attacker
succeeds only if it outputs a single query $q$ for which $\Qry( M,
\aux, q) \neq q(S)$  \emph{subject to the restriction that it did not query $q$ to its $\Qry$ oracle}.
It will be interesting to understand formally the relation between our definition and theirs.
Here are some observations:
\begin{enumerate}
\item Say a scheme is $\epsilon_{NY}$-secure under the NY definition.
Then we claim it is $(t,q, r, q\epsilon_{NY}/r)$-correct. To see
this, assume not. Then there is a $t$-time attacker~$A$ that, with
probability at least $\epsilon'=q\epsilon_{NY}/r$,
issues~$q$ Test-queries of which at least~$r$ cause errors. If we simply run~$A$ and then choose a random Test-query to
output, we succeed with probability at least $\epsilon' \cdot r/q$. It is unclear whether this is tight.

\item There is a contrived scheme that is $(t, q, 2, 0)$-correct but not NY-secure at all; the basic idea is
to have a set-representation that always has \emph{exactly one} false positive that is easy to find
given the public information.

\item If a scheme is $(t,q,1,\epsilon)$-correct, then it is also $\epsilon$-secure under the NY definition,
at least for attackers making at most $q-1$ queries.
This is immediate.
\end{enumerate}

\if{0}
\heading{FP-security of hash-based filters.}
%\tsnote{NY give it in the private hash setting (sort of); but, intuitively, it should be hard to find false-positives even if you know that MD5 or SHA-x are the hash functions.  Not totally sure how to capture this intuition nicely.}
%
%\jnote{Not sure I agree. First of all, if $\epsilon$ is not negligible then a random element is a false positive with noticeable probability. Beyond that, I think that intuitively it makes sense that if you want false positives to be hard to find then you need to make the hash functions secret.}
%
%\tsnote{I don't think it makes sense to talk about asymptotics or negligibility, for exactly the reason you state. I think the key point here is: what is ``hard''? The designer decides what FP rate is acceptable for a given application, a given set of efficiency constraints, the side-effects of a FP, etc.  For a fixed FP target, she optimizes the filter size and/or the number of hash functions for her setting. Whether the target FP rate is 0.01 or $2^{-80}$, we should be able to say (in the ROM, at least) that it will take this-many hash computations, and this many on-line set-membership queries, to find the first FP.  In the standard model, with public hash functions, it is clearly more difficult to make formal claims.  Even so, in practice I expect it would be quite expensive to find a string~$x$ such that $(h_1(x),h_2(x),\ldots,h_k(x))$ falls in a given set of tuples when $h_i(x)=\mathrm{truncate}(\mathrm{SHA256}(\langle i \rangle \concat x), m)$, say. (Even more so if my setting admits relatively expensive hash computations, e.g. SCRYPT.) BTW, your suggestion to have a FP notion with secret~$\mathcal{S}$ is interesting here. }
%
The traditional notion of soundness for a Bloom filter captures the difficulty of finding false positives when the attacker is non-adaptive. Following NY, we consider soundness against adaptive adversaries. But we further generalize to allow for settings in which the attacker may have some a priori knowledge about the contents of~$\calS$. (The Niedermaier et al.~\cite{xxx} attacks, for example, exploit knowledge about the distribution of German last names.) In particular, our notions take as input an explicit distribution $\distr{U}{n}$ over~$[U]^n$.
